---
title: "Peer graded assignment regression models"
author: "yinshu zhang"
date: "October 24, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
\fontsize{8}{8}

```{r setup, include=FALSE}
library(knitr)
library(GGally)
library(ggplot2)
library(leaps)
library(plot3D)
#library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
opts_chunk$set(dev = 'pdf')
options(knitr.table.format = "latex")
attach(mtcars)
mtcars.tab <- read.csv(file = "mtcars.csv", sep = ",")
dim.mtcars <- dim(mtcars)
cor.am <- round(cor(mpg, am),4)
```

# Executive Summary
This study is conducted to explore the relationship between a set of variables and the fuel consumption. The focus of is to address two questions.

> "Is an automatic or manual transmission better for MPG"

> "Quantify the MPG difference between automatic and manual transmissions"

By perform variable F-test and one way ANOVA test, evidences are observed support relationship between transmission type and fule economy. With multi-variable linear regression model, the conclusion is manula transmission results higher MPG number by 2.94 mile per gallon compares to automatic transmission, this result has 1.41 mpg standard error.

# Data Exploration

We will use "mtcars" data to draw conclusions, mtcars is a `r dim.mtcars[1]` by `r dim.mtcars[2]` data frame, see __table 1__ for the description of each column. Each column also been given the *possible* effect to MPG with hifger value, 

Table 1, data columns decription and categories

```{r table, echo=TRUE}
kable(mtcars.tab)
```

Figure 1 in appendex shows the correlation between all variables

# Relatinoship between MPG and AM
## Correlation
From figure 1, we can see the MPG and AM variables has `r cor.am` correlation value, it a indication MPG and AM has relationship, but not strong enough. Therefore, to answer first question, we need better evidences.

We will test the correlation betwen MPG and AM with alternative hypothesis "true correlation is greater than zero"

```{r cortest}
cor.test(mpg, am, alternative = "greater")
```

From the output, p-value is far away from 95% confidence interval, this is a good evidence *supporting* Ha, which is MPG and AM are correlated.

## One way ANOVA 
To further prove the opposite is false, we can use one-way ANOVA. The null hypothesis of the analysis is __the mean MPG are the same between samples__.

```{r anova}
anova1 <- aov(mpg ~ am, data = mtcars)
summary(anova1)
f_crit <- round(qf(.95, 1, 30),2)
```
F value of 16.86 is significiantly larger than the 95% confidence interval critical value `r f_crit`, the result shows strong evident *rejecting* null hypophsis.

# Regression
## simple regression
For the sake of comparsion, let's do a single variable(factor actually) linear model.
```{r reg.sim}
mod.sim <- lm(mpg ~ am, data = mtcars)
summary(mod.sim)
```
To interpurte the output, the coefficient of AM is 7.245, which means changing from auto to manual, will result on average 7.245 mile/gallon fuel consumption increase. However, as expected, the model quality is low, we can't use it to predict the MPG, when only around 36% of residuals varians are explained by this model, AM is a dummy vailable after all.

## seletion of independent variables
To make a multi-variable liner regression model, there are ten variables to choose from. we need to aviod chose highly correlated variables, take a quick look at first summary plot, we can see quite some variables are highly related.

This is not a big surprise, with basic car knowlege we can understand why some of independent variables are correlated. For example, more cyclder of an engine normally means bigger displacement, higher the hourse power normally results lower quater mile time. 

We can categorize those ten variables into following group

1, drivetrain specfications: cyl, disp, drat, vs, carb, drat, am, gear

2, other veichle specfications: wt

3, performance measurements: qsec, hp

we can chose ourselvies and compare the results but better option is to get help from package "leaps".

Here we did a exhausded subset comparsion, with one best result from each number of variable combination. from the right plot, we can see the adjusted R squre does not show big difference after 0.7, the mallow's Cp however have one very low value compares to other combinations. Therefore, we will create regression model with independent variables of weight, quarter mile time, and auto/manual transmission.

```{r multireg}
mod.mul <- lm(mpg ~ wt + qsec + am, data = mtcars)
summary(mod.mul)
```

Interpretation of the output:
This model tells us on every 1000 pounds increase, the veichile fuel economy will reduce by 3.9165 mpg if other two variables remain unchanged. This is expceted as the heavier the car, the wrose the fuel economy. 

 The next line quater mile seconds read as every second increase for a car from stand to reach quater mile, the fule economy will increase 1.2259 mpg, if other variable remain unchanged. The explaination is the lower "qsec" the fast car can travel, normally this is result of more powerful engine, which yeild worse mpg number.

Next line of am is the *answer* to second question, if other variables are unchanged, we only change transmission from auto to manual, it will improve car fuel economy by 2.9358 mpg. 

Now we exame the significiant level of each of those variables. Both weight and quater mile seconds are highly significait to fuel economy, with very small p-value. However the "am" is just below 95% critical value, this tells us transmission type is related to MPG but not as sigificiant as other two variables.



## Regression Residuals
Fianlly we want to check the residual plot and distribution, to verify if 

1, distribution is fairly normal.

2, if multi-variable regression residual has smaller variation.

# Appendix
### Figure 1, variable correlation grid

```{r sumplot, echo=FALSE, fig.dim = c(12, 8), fig.cap="", cache=T }
ggpairs(mtcars)
```

### Figure 2, regression subset selection

```{r subset, fig.dim = c(8,4), echo =F }
regfit.full <- regsubsets(mpg ~ ., data = mtcars, nbest = 1)
par(mfrow=c(1,2))
plot(regfit.full, scale = "adjr2", main="Adj R sqr")
plot(regfit.full, scale = "Cp", main="mallow's Cp")
#mtext("figure 2, regression subset selection", side = 1, outer=T, line = -3)
par(mfrow=c(1,1))
```
### Figure 3, Multi-variable regression model

```{r regmulplot, fig.dim = c(8, 6), fig.cap="", echo=F}
# init grid size
grid.lines = 20
#  fill points in between min and max value of indenpent variables
x.pred <- seq(min(wt),max(wt), length.out = grid.lines)
y.pred <- seq(min(qsec),max(qsec), length.out = grid.lines)
# resize grid
xy <- expand.grid( x = x.pred, y = y.pred)
# fillin z value on, auto and manual will be different grid surface
z.auto.pred <- matrix(
  predict(mod.mul, newdata = data.frame( wt = xy$x, qsec = xy$y, am=0)), 
  nrow =grid.lines, ncol = grid.lines
  )
z.manual.pred <- matrix(
  predict(mod.mul, newdata = data.frame( wt = xy$x, qsec = xy$y, am=1)), 
  nrow =grid.lines, ncol = grid.lines
  )
# plot auto and manual individually, later plot will overlapping first
points.auto <- mtcars[mtcars$am == 0,]
points.manual <- mtcars[mtcars$am == 1,]

scatter3D(points.auto$wt, points.auto$qsec, points.auto$mpg, 
          ylim = c(10,30), zlim = c(0,40), 
          col = "blue", ticktype = "detailed",
          colkey=F, pch=20, cex=2, alpha = 0.5,
          phi = -5, theta = 40, bty = "g", 
          xlab ="wt", ylab="qsec",zlab="mpg",
          surf = list(x = x.pred, y = y.pred, z=z.auto.pred, facets = NA , col = "blue"),
          main = "finger 3, MPG with auto(blue) and manual(red)"
)
scatter3D(points.manual$wt, points.manual$qsec, points.manual$mpg, 
          col = "red", alpha = 0.5,
          pch=20, cex=2 , 
          surf = list(x = x.pred, y = y.pred, z=z.manual.pred, facets = NA , col = "red"),
          add = T
)
```

### Figure 4, regression residuals

```{r residualcomp, fig.dim = c(8, 6), fig.cap="" , echo = F}
se <- function(x) sqrt(var(x)/length(x))
par(mfrow=c(2,2))
res.range <- ceiling(max(sapply(list(mod.sim$residuals, mod.mul$residuals), range)))
plot(mod.sim$residuals, ylim = res.range * c(1,-1), main="simple", ylab = "variation")
abline(h=0)
plot(mod.mul$residuals, ylim = res.range * c(1,-1), main="multi", ylab = "variation")
abline(h=0)
hist(mod.sim$residuals, breaks = 30, main="simple", xlab = "variation")
hist(mod.mul$residuals, breaks = 30, main="multi", xlab = "variation")
mtext("figure 4, residuals", side = 1, outer = T, line = -3)
par(mfrow=c(1,1))
```
